Cryptanalysis of SBC based post-quantum signature scheme

### 2025-06-16 rapport.tex
minor changes

	\section{MPCitH and SBC signature scheme}
		This section presents the MPCitH framework, the Subfield Bilinear Collision (SBC) problem, and the associated signature scheme.

		\subsection{Multiparty Computation}

		A standard example is to make construct a $(N-1)$-private arithmetic circuit among $N$ participants and they all broadcast their results. See figure \ref{fig:mpc}.

		\begin{figure}[h]
			\centering
			\begin{tikzpicture}[
				participant/.style={
					thick, 
					minimum size=2cm, 
					inner sep=0pt
				},
				image/.style={
					anchor=center
				},
				arrow/.style={
					-{Latex[length=3mm, width=2mm]}, 
					thick
				}
				]

				% Nœuds des participants avec images
				\node[participant] (A) at (0,0) {\includegraphics[width=\participantimgsize]{flash.png}};
				\node[participant] (B) at (2,-3) {\includegraphics[width=\participantimgsize]{hudson.png}};
				\node[participant] (C) at (5,-3) {\includegraphics[width=\participantimgsize]{martin.png}};
				\node[participant] (D) at (7,0) {\includegraphics[width=\participantimgsize]{king.png}};
				\node[participant] (E) at (2,3) {\includegraphics[width=\participantimgsize]{sally.png}};
				\node[participant] (F) at (5,3) {\includegraphics[width=\participantimgsize]{sheriff.png}};

				% Communications (flèches)
				\draw[<->] (A) -- (B);
				\draw[<->] (A) -- (C);
				\draw[<->] (A) -- (D);
				\draw[<->] (A) -- (E);
				\draw[<->] (A) -- (F);
				\draw[<->] (B) -- (C);
				\draw[<->] (B) -- (D);
				\draw[<->] (B) -- (E);
				\draw[<->] (B) -- (F);
				\draw[<->] (C) -- (D);
				\draw[<->] (C) -- (E);
				\draw[<->] (C) -- (F);
				\draw[<->] (D) -- (E);
				\draw[<->] (D) -- (F);
				\draw[<->] (E) -- (F);

			\end{tikzpicture}
			\caption{MPC protocol with broadcast communication}
			\label{fig:mpc}
		\end{figure}

		To evaluate the circuit that depends on the secret witness $w$, each party holds a secret share $[w]$. They perform local computations on their respective shares, then broadcast the results, which are subsequently combined to obtain the final output.

@@

		Since the introduction of the framework, many optimizations have been published to reduce the communication or the computation cost. We will discuss some of them.

@@

		In our case, we do not transmit the seed due to the use of linear secret sharing. Since we typically employ an $(N-1)$-private MPC protocol, we only share the co-path of the $i$-th leaf, which remains hidden from $\mathcal{V}$. As a result, instead of sharing $N$ values, we only need to share $log_2(N)$ values.

		\subsubsection{HyperCube technique}

@@

		\subsection{NSBC variant}

		The authors also introduce a particular version of this problem called normalized SBC (NSBC) where $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ and
		$\vec{x} = (\vec{x'}, 1, 0), \vec{y} = (\vec{y'}, 0, 1)$ to ensure that they are not colinear without needing to regenerate each time they are not. We also denote $(\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ if $(\vec{x}, \vec{y})$ are solutions to the NSBC problem on $(\vec{u}, \vec{v})$.

@@

		The signature protocol is described in .. with the signature and verification algorithms.

@@

	\subsection{Attacks summary}
		\begin{table}[h!]
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				Algorithm & Parameters & Time & Memory \\
				\hline
				Exhaustive search & / & 260 & 24 \\
				\hline
				Dinur First & / & 229 & 200 \\
				\hline
				Dinur Second & / & 221 & 172 \\
				\hline
				Crossbred & D=12, d=10, k=130 & 159 & 90 \\
				\hline
				Naive & / & 150 & 15 \\
				\hline
			\end{tabular}
			\caption{Time and memory complexity of given algorithms in powers of 2.
				For Exhaustive search, Dinur First and Dinur second, the estimations were obtained using CryptographicEstimator \cite{EVZ+24}}
		\end{center}
		\end{table}

>>>>

	\section{MPCitH and SBC signature scheme}

@@

		We introduce some standard notations in MPCitH protocols. 
		\begin{itemize}
			\item $[w]$, $w$ is a shared value.
			\item $w^{\llbracket i \rrbracket}$ is the $i$-th share of w.
			\item $w^[e]$ is a value used in the $e$-th round of the protocol.
			\item $\delta_x$ is the offset for the additive sharing of $x$ s.t. $x = \delta_x + \sum_i x^{\llbracket i \rrbracket}$
		\end{itemize}

		Finally, we discuss several optimizations that have been proposed in the literature since and are now considered standard and widely used.

@@

		In our case, we do not transmit the seed due to the use of linear secret sharing. Since we typically employ an $(N-1)$-private MPC protocol, we only share the co-path of the $i$-th leaf, which remains hidden from $\mathcal{V}$. As a result, instead of sharing $N$ values, we only need to share $log(N)$ values.

		\subsubsection{HyperCube technique}
 		This optimization technique was introduced in \cite{AGHH+22} and is commonly used in MPCitH schemes. This geometric approach models the $N$ shares as forming a $d$-dimensional hypercube of side length $k$. The often used parameters are $k = 2$ and $N = 2^d$, so the number party counts reduces to $log(N) + 1$. This technique often leads to MPC computations less expensive, smaller proof sizes and larger party sizes. However, these benefits come at the cost of increased reliance on GGM tree generation, which can become computationally expensive.

@@

		\textbf{NSBC variant}.
		The authors also introduce a particular version of this problem called normalized SBC (NSBC) where $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ and
		$\vec{x} = (\vec{x'}, 1, 0), \vec{y} = (\vec{y'}, 0, 1)$ to ensure that they are not colinear without needing to regenerate each time they are not. We also denote $(\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ if $(\vec{x}, \vec{y})$ are solutions to the NSBC problem on $(\vec{u}, \vec{v})$.

@@

		\subsection{Multiparty protocol for the SBC problem}

		\subsection{Identification Protocol}
		We present the identification protocol, which serves as a basis for constructing the signature scheme using the Fiat-Shamir transformation.

@@

	\subsection{Attacks summary}
		\begin{table}[h!]
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				Algorithm & Parameters & Time & Memory \\
				\hline
				Exhaustive search & / & 260 & 24 \\
				\hline
				Dinur First & / & 229 & 200 \\
				\hline
				Dinur Second & / & 221 & 172 \\
				\hline
				Crossbred & D=12, d=10, k=130 & 159 & 90 \\
				\hline
				Naive & / & 150 & 15 \\
				\hline
			\end{tabular}
			\caption{Time and memory complexity of given algorithms in powers of 2.
				For Exhaustive search, Dinur First and Dinur second, the estimations were obtained using CryptographicEstimator \cite{EVZ+24}}
			\label{tab_attacks}
		\end{center}
		\end{table}
